\chapter{Formulating a research question}
\label{ch:researchquestion}

This chapter provides some suggestions on searching for a topic. The website of the HOGENT library also has some general advice about this\footnote{\url{https://bib.hogent.be/how-to/onderwerp-formuleren/inleiding/}}. This guide is specifically aimed at the bachelor applied Informatics though.

The ultimate goal of the bachelor's thesis is to prove yourself capable of delivering substantiated advice on the application of state-of-the-art IT technologies within a professional context. 

This means you have to be able to gather and process information from different sources with a critical mindset. Your first source of information will usually be the target audience of your thesis, for example when compiling a list of requirements. Additionally, you need to be able to search through professional literature and judge whether sources are reliable and relevant for your particular use case. Ultimately, you will find yourself gaining new knowledge, usually in the form of a conclusion based on test results or a software prototype.

A good research subject for a bachelor's thesis should be a \emph{specific problem within a concrete use case or scenario}, requiring you to find the best possible solution (for this case!) given the current state of technology. The target audience is very specific, f.e. a single company (team).

\section{Types of research}
\label{sec:researchtypes}

A bachelor's thesis does not have to try to imitate an academic master's thesis. The professional bachelor has a different value, so it is perfectly possible to excel within the specific uniqueness of this profile. The main difference lies mainly in the type of research that is conducted.

In \emph{fundamental research}, which is typically carried out at universities, the emphasis is on expanding our knowledge in general. For example, within computer science it can be about developing new and/or more efficient algorithms for problem solving. Whether the results of fundamental research are immediately applicable is of secondary importance.

\emph{Applied research}, on the other hand, tries to formulate an answer to a concrete question from the field. The researcher will then try to answer that question on the basis of the knowledge that is currently available (and published by experts on the subject). This may concern, how a new technology can be applied concretely in a business context, making a choice between various alternative products or technologies, a preliminary research prior to developing an application, etc. In applied research, the target group is also very specific, for example a single company. We also notice that the topics introduced by companies are also the best developed and most often lead to a good bachelor's thesis.

Applied research is in no way inferior to fundamental research. Rather, both are complementary. Results of fundamental research are often not applicable in practical use cases. They might be based on \textit{toy problems} or the research question might be too abstract to reflect real-life situations. This is why applied research is required: to make sure the results of fundamental research are put into practice to solve actual problems in the field.

\section{Research domain choice}
\label{sec:research}

The first step is to choose your research domain. This is something that no one can really help you with. Choose a domain that you are interested in, so that you can gain sufficient motivation to delve into it. Your chosen specialization in the final year may be a good starting point. What kind of job would you like to start after graduation? Which technologies, platforms {\ldots} would you like to work with?

Search for current problems/challenges in your chosen field. It is important to take your time to ``immerse'' yourself in current events. This cannot be done overnight. It is more efficient to put some time into this regularly over a number of weeks (e.g.\ an hour every day). In the long run you should be able to recognize the most important themes that people are mainly working on at the moment and that could provide inspiration for your subject.

In Section~\ref{sec:searchingforrelevantinformation} you will find some concrete tips and starting points on searching for relevant information. 

\section{Formulating research questions and targets}
\label{sec:formulatingresearchquestion}

Once you get a feel for recent changes on a topic, you typically also get to know the most important problems and points of discussion. These can lead to the formulation of your main research question, which you can further divide into more concrete research sub-questions.

A good research question for a bachelor's thesis starts from a real or at least realistic \textit{company case}, a concrete problem that a specific company struggles with. This immediately ensures a clear \textit{delimitation} of the subject. A research question that is too vague or general will never lead to a good bachelor's thesis.

For example, ``What is the best agile methodology for companies?'' is \textit{not} a good research question in that regard. There are several agile methodologies, and there's probably a good reason for that. There is no clear winner that is best suited for every business. If that were the case, then twenty years after the publication of the Agile Manifesto \autocite{BeckEtAl2001}, it would already be clear what the best methodology is. A suitable methodology depends on many factors: the company culture, the type of company, the type of IT solutions being developed, the size of the company, etc. If you formulate your research question too general, you will never be able to come to a conclusive conclusion. We have experienced this too often in the past. The student then writes in the conclusion ``The best solution depends on your preference and your specific situation.'' In other words, there is no conclusion. So \textbf{start from a concrete case}, e.g. company XYZ struggles with the delivery of software on time and within budget and wants to apply an agile methodology. What is the most appropriate methodology for their specific situation? This clearly defines the subject and its added value is immediately visible.

There is no conclusive answer to a good research question yet. Questions such as ``What is data mining?'', ``Which PHP frameworks are there?'', ``What is the history of the Cloud?'' are therefore not suitable, because the answer can be found quickly by search on Wikipedia or Google, or by checking relevant professional literature.

It is also important to immediately think about the expected result, the \textit{research objective}. Under what circumstances can you speak of a successful bachelor's thesis? What is the concrete end result that you want to achieve?

You can use the S.M.A.R.T. principle of \autocite{UchelenJungjohann2003} to formulate your research objective(s):

\begin{description}
    \item[Specific] Is the objective unambiguous? Is it clear who will benefit?
    \item[Measurable] Under which (measurable/observable) conditions or form has the goal been achieved? Is the goal to deliver a proof-of-concept or prototype? An analysis for an application to be developed? A comparison between possible alternative solutions with clear advice? etc.
    \item[Acceptable] Are these goals acceptable to the target audience? Will the target group actually be able to use the proposed solution?
    \item[Realistic] Is the goal achievable? Is the subject sufficiently defined? Is the difficulty of the problem at the professional bachelor level?
    \item[Time-bound] When (in time) should the goal be reached? This is clearly laid down for a bachelor's thesis: the deadline for submission. 
\end{description}

\section{Types of bachelor theses}
\label{sec:types_of_theses}
In this section, we list some thesis topic types often chosen by students. They can serve as inspiration for your own subject. Some types of topics are rather specific to a specialisation, others are possible within different specialisations.

\subsection{Comparative study}
\label{ssec:thesis_type_comparative_study}

You can basically work out this kind of topic for any specialisation. 

Starting from a specific case or problem situation, you look for possible
solutions. You start with a requirements analysis to gather your requirements and prioritise them (f.e. using the MoSCoW method). You then browse professional literature for possible solutions and collect these in a long list. On the basis of publicly available information, you check them against the requirements and keep the most promising in a short list. You then build a proof-of-concept (PoC) showing the chosen solution is feasible (or not), highlighting any unresolved issues that still need further research.

Example (in Dutch):

\begin{itemize}
	\item \emph{System \& Network Engineering:} \citetitle{VanDamme2021} \autocite{VanDamme2021}.
\end{itemize}

\subsection{Business case}%
\label{ssec:thesis_type_business_case}

You can basically work out this kind of topic for any specialisation.

A company is considering using an innovative technology, but this requires an investment in time, resources and expertise without any guarantee of worthwhile results. You are tasked with preparing a business case to identify the costs and benefits of this investment. A critical and objective view is essential! Don't be swayed by hype, marketing material or statements from suppliers or consultants.

You start with a description/inventory of the current situation and any problems or pain points needing a solution. You collect requirements that must be taken into account when applying the new technology. You study the technology  at hand and aim to explain exactly how it works, down to the most technical details (applied to the use case).

You build a proof-of-concept to validate to what extent the new technology can provide a solution to existing problems. Based on this, you formulate a recommendation and summarise the costs and benefits.

Examples (in Dutch):

\begin{itemize}
	\item \emph{Mainframe Expert:} \citetitle{Coucke2020}} \autocite{Coucke2020}.
	\item \emph{System \& Network Engineering:} \citetitle{Vanduynslager2024} \autocite{Vanduynslager2024}}
\end{itemize}

\subsection{Preliminary research/analysis of an application to be developed}%
\label{ssec:thesis_type_application_analysis}

You can basically work out this kind of topic for any specialisation.

This type of topic is typical of the specialisations \emph{Functional \& Business Analysis} and \emph{Mobile \& Enterprise Development}. You have an idea for a specific application that you want to develop, or that a company/organisation has a need for. You start with gathering requirements and prioritising them (e.g. using the MoSCoW method). You formulate use cases and define a minimal viable product. You identify technologies that are eligible to be used in the application. This could involve hardware, sensors, programming languages, frameworks, databases or storage {\ldots}

You make well-informed choices about the architecture of the application: which components will the application consist of, what will the backend/frontend look like {\ldots} You can build a prototype or proof-of-concept to demonstrate that the chosen technologies and architecture work. Your research might serve as a starting point for actual development later on.

Example (in Dutch):

\begin{itemize}
	\item \citetitle*{VanDerPlaetsen2013} \autocite{VanDerPlaetsen2013}.
\end{itemize}

\subsection{Cybersecurity risk analysis}%
\label{ssec:thesis_type_cybersecurity_risk_analysis}

This kind of topic obviously fits within the specialisation of \emph{System \& Network En-
gineering} but could also fit under \emph{Mobile \& Enterprise development} when screening a specific web or mobile application.

You examine a specific device, application, system, network {$\ldots$} for possible vulnerabilities. You start by describing the existing
situation. How is it being used? Which data is being stored and in which manner? Which security measures are already in place? You identify
potential vulnerabilities and threats. You often build a test setup for
demonstration. You estimate the probability and impact of the threat(s) since the product of both results in the risk you're facing. You can then
prioritise these risks and formulate recommendations to address, reduce, avoid or even accept them.

Example (in Dutch):

\begin{itemize}
	\item \citetitle*{Vermeersch2024} \autocite{Vermeersch2024}.
\end{itemize}

\subsection{Application of AI/ML}%
\label{ssec:thesis_type_application_of_AI_ML}

This kind of topic is often chosen by students within the specialisation of \emph{AI \& Data Engineering}.

You want to use an existing Artificial Intelligence or Machine Learning technique to solve a specific problem. You describe the problem and formulate the concrete learning task. 
\begin{itemize}
	\item What exactly should the system be able to do?
	\item What kind of data do you need to train this system? Describe this in as much detail as possible!
	\begin{itemize}
		\item What is needed to obtain or generate this data? 
		\item Do you have access to the necessary data, either a public dataset or through the company or organisation that wants to use the system?
		\item Do you need to collect or generate it yourself? If so, how? 
	\end{itemize}
	\item How would the trained model function \textit{in production}? On what kind of computer system? Server, desktop, laptop, mobile phone, tablet, a single board system such as a Raspberry Pi {\ldots}?
	\item How is input being fed to the system (logs, sensors, camera {\ldots}) and what should the output look like?
\end{itemize} 


Search the scientific literature for review papers where similar problems have already been investigated by applying certain techniques. Which technique seems most suitable for your problem? Select at least one technique in advance for you to implement. Build a proof-of-concept to demonstrate that the chosen technique works for this problem. Apply an iterative, agile methodology to systematically improve your system until you get usable results.

Example (in Dutch):

\begin{itemize}
	\item \emph{AI \& Data Engineering:} \citetitle*{Cluyse2023} \autocite{Cluyse2023}.
\end{itemize}

\section{Writing a research proposal}
\label{sec:researchproposal}

Once you have a research question, you can write out your topic to submit it for approval. This means that you will also be conducting some literature research. For instructions on how to approach this, see Chapter~\ref{ch:literaturestudy}.

What you read about the subject, you have to structure and formulate in your own words in a continuous text. A good tool to organize your thoughts on a topic and to write a structured text around it later is to set up a mind map. There are various tools that you can use for free, such as XMind\footnote{\url{https://www.xmind.net/}} or FreeMind\footnote{\url{http://freemind.sourceforge.net/ }}.

% TODO: voorbeeld mindmap?

It is not the intention at this stage to write a fully detailed literature study, but a number of references are expected. After reading your proposal, your supervisor should understand the context of your research and why there is a problem that needs a solution. You must be able to demonstrate this on the basis of authoritative professional literature.

A good literature review should contain two main aspects. On one hand you have to address the \textbf{problem domain}, meaning you should explain the case or situation at hand. The nature of the problem and any required context should be discussed in this part of your text.

On the other hand, you will have to explore the \textbf{solution domain}, touching upon the different possible solutions which have proven effective in similar situations.  This section should prove you already know the general direction in which to look for a solution.

\section{Formulating a title}
\label{sec:formulating_a_title}

Don't postpone thinking about a proper title for your bachelor's thesis. It doesn't have to be final yet, but a good title makes it clear exactly which direction you want to go. A concrete title gives your supervisor(s) the confidence that you know exactly what you want to do and that this is a realistic goal. Try to keep the following in mind when formulating a title:

\begin{itemize}
    \item Do not formulate the title as a question.
    \item Avoid obscure jargon and do not use abbreviations. The title should also be understandable to someone outside your specific field.
    \item Just naming your professional domain is not enough because it is too vague. Your title should be specific and make it clear what exactly you want to research. That also means that a title can be long. For example, ``Cloud computing'' is a general term that covers a broad range of topics, and thus means nothing. ``The selection of an open source Infrastructure as a Service platform for setting up a test environment for web development'' is a lot more concrete.  
\end{itemize}

When assessing a subject, the following criteria are taken into account:

\begin{itemize}
 \item There is a concrete, \textbf{clearly defined research question}, research objective.
 \item The proposed subject  starts from \textbf{an actual problem}, which can be described in detail.
 \item The proposal has a \textbf{clear and specific target audience} within the ICT field.
 \item The research can only be carried out by someone skilled in applied IT. This means the subject has a clear \textbf{technical and technological component}.
 \item The \textbf{methodology} is clearly justified, research techniques are suitable for answering the research question
 \item  There is a \textbf{practical component}. The research is not limited to a literature review.
 \item There is a clear \textbf{personal contribution} regarding a \textbf{research component}. No a priori choices have been made. 

\end{itemize}

\section{Common mistakes}
\label{sec:subjectcommonmistakes}

The following things provide us with the idea that your subject has not yet been sufficiently worked out in depth or that there still are important mistakes that stand in the way of a successful bachelor's thesis:

\begin{itemize}
  \item There is no concrete real or realistic \textbf{company case} associated with the topic. Research at bachelor level is applied, in other words we try to solve concrete, real problems. That should also be reflected in your subject.
  \item Your \textbf{target audience} is not specific enough. Who will benefit from your solution? Who should be interested in your conclusion? ''Mobile developers'' or ''Database administrators'' for example are too vague an audience.
  \item (One of) the research objective(s) is \textbf{speculation about the future} (e.g.  ``what will be the impact of the Internet of Things on everyday life?''). The conclusion of a bachelor's thesis must be verifiable, future predictions never are.
  \item The result of the research is \textbf{depending on external factors} to a large extent. For example, if you are going to conduct a survey, it is important to realize that it is not easy to find enough respondents. You will therefore immediately have to indicate how you intend to take a sufficiently large sample. Even if your plan is to conduct interviews with experts, companies, etc., it is important to make the necessary contacts before the start of your research. After all, if it is not possible to speak to the necessary people in time, the result of your bachelor's thesis will be in danger.
  \item The subject shows insufficient \textbf{technical complexity}. The professional bachelor's degree in applied computer science is a technical profile and we expect you to demonstrate that you are technically strong. For example, a topic such as ``what is the most user-friendly mobile application for home banking for seniors?'' \textit{not} meets this requirement. These kinds of applications are developed by banks and as a student you will never have access to the internal workings. You may have to limit yourself to interviews and/or surveys with the target group. In this way you cannot demonstrate what you have to offer as a computer scientist.
  \item The \textbf{reference list} is too short or consists of unsuitable sources. Take the time to do an initial literature review and look beyond the first blog entry. In chapter~\ref{ch:literaturestudy} you will find instructions on how to approach this.
  \item The subject is \textbf{outside the scope of a bachelor's program or of the student}. Choose a subject that corresponds to a bachelor's degree in terms of level. A typical example of a subject that is too difficult is \textit{Quantum Computing}, at least for a bachelor's degree. Such a subject may be eye-catching, but you are taking the risk of shooting yourself in the foot in terms of difficulty and practical feasibility.     
\end{itemize}

\section{Summary}
\label{sec:subjectsummary}

\begin{itemize}
 \item Take your time to find a topic that really interests you.
 \item When writing up a proposal, it is best to be as \emph{concrete} as possible.
 \item Avoid common mistakes such as the lack of a concrete case, speculation, dependence on external factors or a literature study that is too superficial.
\end{itemize}

The most important criteria used during evaluation of a subject are:

\begin{itemize}
	\item The title
	\begin{itemize}
		\item The (concrete) research questions can be deduced from the title (and this is not a yes/no-question).
		\item The title is a statement, not a question.
		\item The title does not contain overly technical terms or abbreviations.
	\end{itemize}
	\item Suitability of the topic
	\begin{itemize}
		\item The proposal includes a clear technical and practical component.
		\item There is a clear personal contribution on part of the student.
		\item The proposal is S.M.A.R.T.
	\end{itemize}
	\item Delineation of the topic
	\begin{itemize}
		\item The topic is sufficiently narrowed down, concrete and specific.
		\item The proposal contains a concrete, specific problem with relevant sub-questions.
		\item The target audience is well-defined and very specific.
	\end{itemize}
	
\end{itemize}
